An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}

### 2022-09-23 paper/paper1/ms.tex
neat discussion

@@

\subsection{Identifying missing lines}\label{secMissingLines}
The automatic abundance determination spirit of  \emph{blas\'e} resembles \texttt{ROBOSPECT} \citep{2013PASP..125.1164W}, and other approaches aimed at automatically measuring equivalent widths for a large number of stars.  One missing piece here is either to post-process the  \emph{blas\'e}-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of  \emph{blas\'e}.  The latter case is intriguing.  For example, PHOENIX may not yet be aware of the existence of some obscure lanthanum (La) line, possibly because the line was not yet understood and incorporated at the time of PHOENIX computation.  Rather than wait for PHOENIX to receive the necessary update,  \emph{blas\'e} could instantiate one or more such lines in the locations guided by the atomic line list.  This line could get initialized with some small amplitude, and would get added onto the underlying PHOENIX source as yet another freely tunable line.  In this way, all the other nearby lines would be used for assigning continuum and blending, but this free-standing lanthanum line would acquire a value properly taking into account the impact of its neighbors.

>>>>

\edit1{\emph{``Fancy interpolator''}--  A linear interpolator has to store the slope, offsets, and start and end points of all the line segments from pixel-to-pixel, merely to recreate a spectrum losslessly at some arbitrary coordinate or coordinates.  \emph{Blas\'e} lossily recreates a spectrum, while having to store less information.  In this way, it is fair to think of \emph{Blas\'e} as an interpolator that happens to have an intuitive interpretation.}

@@

\subsection{Identifying missing lines}\label{secMissingLines}
\edit1{So far we have not addressed the inevitable prospect that some spectral line is present in the data, but was not predicted at the time of the atmosphere simulation.  This ``missing lines'' scenario arises from our incomplete and evolving knowledge of quantum mechanics and molecular chemistry.  \texttt{ROBOSPECT} \citep{2013PASP..125.1164W} and the \emph{Cannon} \citep{2017ApJ...836....5H} have data-driven ways to identify these lines.  The \texttt{Cannon} can go further to indicate how these previously unknown lines correlate with the stellar labels, providing some physical undersanding for their cause. \emph{Blas\'e} as it currently stands has no mechanism for identifying these lines.  Here we consider two scenarios: \emph{I)} information about the lines now exists, but they were not included in the spectrum forward modeling process, or \emph{II)} the lines are truly anomalous and unknown.  For the first scenario, a Voigt profile or profiles could be instantiated on top of the line center location(s) of the previously unknown line, and it would be fitted alongside all its other fellow lines:}

\begin{eqnarray}
    \mathsf{S}_{\rm clone} = \Big( \prod_{m=1}^{N_{\mathrm{missing}}} 1-a_m \mathsf{V}_m \Big) \Big(\prod_{j=1}^{N_{\mathrm{lines}}} 1-a_j \mathsf{V}_j \Big) \label{missing}
\end{eqnarray}

for a large number of stars.  One missing piece here is either to post-process the  \emph{blas\'e}-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of  \emph{blas\'e}.  The latter case is intriguing.    Rather than wait for PHOENIX to receive the necessary update,  \emph{blas\'e} could instantiate one or more such lines in the locations guided by the atomic line list.  This line could get initialized with some small amplitude, and would get added onto the underlying PHOENIX source as yet another freely tunable line.  In this way, all the other nearby lines would be used for assigning continuum and blending, but this free-standing lanthanum line would acquire a value properly taking into account the impact of its neighbors.
